Supplemental Materials: \\ \emph{Surface-driven registration method for the structure-informed segmentation of diffusion MR images}

### 2016-02-02 Publications/2015-NeuroImage/04-methods.tex
added fixes of all coauthors, and spell check

\begin{figure}
  \includegraphics[width=\linewidth]{figure01}
  \caption{The {\color{blue} interfacing surfaces $\Gamma_m$ between the competing \glspl{roi} $\Omega_l$,
  play the role of active contours which drive the registration process.}
  They evolve iteratively following {\color{blue} the normal $\hat{\mathbf{n}}_i$ of the surface} at each vertex
    $\vec{v}_i$ of the mesh.
  The gradient speeds $\bar{s}_i$ drive registration, which are computed as the disparity of the data
    energies with respect to the two limiting regions of $M(\vec{v}_i)$, the features of the image
    $M$ in the location of vertex $\vec{v}_i$.
  {\color{blue} The computation of shape-gradients is developed in \ref{app:shape_gradients}.}
  In this figure, the $\bar{s}_1$ derivated from \autoref{eq:regseg-shape_gradient_disc2} is 
    written in the lower box, with $\Omega_{wm}$ in the inner limiting region, $\Omega_{gm}$
    the outer region, and $w_{0,1}$ is the relative area associated with vertex $\vec{v}_1$
    with respect to the total area of surface $\Gamma_0$.
  }\label{fig:regseg-method}
\end{figure}

\begin{figure*}
% \includestandalone[width=\linewidth]{figure02-model}
\includegraphics[width=\linewidth]{figure02}
% \includegraphics[width=\linewidth]{figure04B}
% \includegraphics[width=\linewidth]{figure04C}
\caption{%
Evolution of the segmentation model defined by the homogeneous regions $\Omega_l$,
  for one real dataset.
{\color{blue}%
Panel (A, left) shows the joint distribution of the \gls*{fa} and \gls*{adc} conditioned to
  the segmentation \omegaset{} defined by the surfaces $\gammaset_R$ extracted from
  the \gls*{t1} image.
The plot was generated for reference using undistorted diffusion data, and
  therefore, $\gammaset_R$ is aligned with the \gls*{fa} and the \gls*{adc}.
The problem arises when the diffusion data present deformation, and the
  contours $\gammaset_R$ do not fit within the data (A, center).
After registration with \regseg{}, the contours are mapped onto the
  diffusion data (A, right), and the joint density plot is closer to the reference
  situation.
In panel (B), the three plots in (A) are decomposed tissue-wise.
Using filled contours, the bivariate distribution of each tissue is highlighted in its designated color,
  and represented over the remaining tissues (in gray colors).
To help assessment, dashed contours in black-to-white colors represent
  the corresponding distribution in the reference plot.
The registration process optimizes the segmentation model of \regseg{},
  and thus, the distribution of each region after registration is located
  closer to that corresponding in the reference situation, the shape of the distribution
  is more similar to the reference, and their spread is also reduced.
The effects of optimization are more noticeable on the \gls*{gm} ($\Omega_\text{GM}$) and
  the \gls*{wm} ($\Omega_\text{WM}$).
Particularly, the \gls*{wm} typically shows a bimodal distribution when the contours
  \gammaset{} do not fit the data.}
% The left plot in each panel represents a kernel density estimate of the location and spread of
%   each region $\Omega_l$ in the bivariate feature space of $M$, which comprises
%   the \gls*{fa} and \gls*{adc} maps.
% At the top and the right margins, the marginal distributions of $\Omega_l$ on the \gls*{fa} and
%   \gls*{adc} are respectively displayed.
% On the right side, the sampled distribution of each region is shown in small scatter plot, in contrast to
%   the complete bivariate data.
% A) shows the initialization of the model, when the anatomically-correct surfaces are not
%   mapped into the \gls*{dmri} space.
% The panel shows the resulting conditional distributions of segmenting the (distorted) \gls*{fa} and
%   \gls*{adc} maps with the original (undistorted) surfaces.
% Panel B shows the equivalent information, using the surfaces that are perfectly mapped into the
%   distorted \gls*{dmri} space.
% Since the surfaces are now aligned in $M$, the spreads of the conditional distributions for each region are
%   reduced with respect to those in A.
% Panel C, after \regseg{} registration, shows how the covariance matrices $\boldsymbol{\Sigma}_l$
%   are minimized with respect to panel A, and the resulting segmentation is close
%   to the theoretical situation presented in panel B.}
}\label{fig:regseg-model}
\end{figure*}

@@

\revcomment[R\#3-C.20]{%
  \begin{align}
  \frac{\partial E(\vec{u})}{\partial \vec{u}_k} &=
  \frac{ \partial }{\partial \vec{u}_k} \Big\{
  \underset{l}{\sum} \big[ \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \big] \,
  + \int_{\Omega} \frac12 [ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
  + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} ] \,d\vec{r}
  \Big\},
  \label{eq:regseg-gradient_descent}
  \end{align}}
  where $\vec{u}^{\circ2} = \vec{u}^T \cdot \vec{u}$, \revcomment[R\#3-C.21]{%
  and $\{\boldsymbol{\alpha}, \boldsymbol{\beta}\}$ are the expected variances along
  the imaging axes of the displacements field and its gradient, respectively.}.
Then, the data and regularization terms are split and discretized to compute their
  derivatives.
\revcomment[R\#3-C.14]{%
The derivative of the data term is computed using explicit shape gradients (see \ref{app:shape_gradients}),
  that finally lead to obtain vertex-wise speeds of the gradient $\bar{s}_i$ as}
  illustrated in \autoref{fig:regseg-method}.
\revcomment[R\#3-C.14]{%
Using the expressions \eqref{eq:regseg-gradient_wshape} and 
  \eqref{eq:regseg-shape_gradient_final} given in \ref{app:shape_gradients}
  and introducing the analytical derivative of the regularization term, then
  the \autoref{eq:regseg-gradient_descent} is reformulated as:}

  \begin{align}
  \frac{\partial E(\vec{u})}{\partial \vec{u}_k} &=
  \vec{g}_k  + \boldsymbol{\alpha} \cdot \vec{u}_k - \boldsymbol{\beta} \cdot (\Delta \vec{u}_k).
  \label{eq:regseg-final_gradient}
  \end{align}

Finally, to descend this gradient, we establish a semi-implicit Euler scheme
  \citepalias[see][section S1.3]{esteban_useful_2015},
  with a step size parameter $\delta$, which we solve in the spectral domain as follows:

@@

\paragraph*{Implementation details and convergence}%
\label{sec:regseg-conv_report}
The \regseg{} tool includes a multiresolution strategy on the free-form deformation field.
Registration pyramids are created by setting the spacing between the control points of the B-spline basis
  functions for each level of the multiresolution strategy.
The implementation details as well as other features (such as the sparse matrix approach
  to fast interpolation) and the main parameters
  (such as $\delta$, $\boldsymbol{\alpha}$, $\boldsymbol{\beta}$, the B-spline grid resolutions,
  and target image smoothing) are discussed in the \citetalias[][section S1]{esteban_useful_2015}.

@@

\begin{figure*}
\includegraphics[width=\linewidth]{figure03}
\caption{Evaluation of \regseg{} using phantom data according to the following instrumental workflow.
  1) The reference surfaces $\gammaset_R$ are triangularized meshes extracted from the four binary shapes (i.e., ``box'', ``ball'', ``L'', ``gyrus'').
  2) A ground-truth displacement field was generated as described in \autoref{sec:regseg-digital_phantoms}, and applied to warp
      $\gammaset_R$, thereby obtaining $\gammaset_\text{true}$.
  3) After being warped, $\gammaset_\text{true}$ were projected onto the corresponding discrete 3D volume and downsampled to create partial volume effects at two resolutions,
     i.e., \isores[mm]{2.0} and \isores[mm]{1.0}, thereby producing sets of tissue fractions maps.
  4) The tissue fractions were fed into a \acrfull*{mr} simulator, which generated \acrfull*{t1} and \acrfull*{t2} -like images at the
     two possible resolutions.
  5) The \regseg{} tool was applied using the warped test images as multispectral moving images and $\gammaset_R$ as shape priors.
  6) The agreement between the surfaces fitted by \regseg{} ($\hat{\gammaset}_\text{test}$) and $\gammaset_\text{true}$ were assessed
     visually using automatically generated visual reports and quantitatively with the Hausdorff distance between the
     corresponding surfaces.}\label{fig:regseg-evphantoms}
\end{figure*}

@@

\paragraph*{Settings} %
\Regseg{} accepts an affine mapping from surface-space to the \gls*{dmri} data as initialization.
However, the images provided by the \gls*{hcp} are already spatially normalized.
Therefore, the initial estimation of distortion is zero in this experiment.
Since the distortion $U_\text{true}$ is aligned along the \gls*{pe} direction ($y$-axis in our settings),
  \regseg{} was configured to allow nonzero displacements only on that corresponding direction.
For the experiments on real data, \regseg{} established a multi-resolution pyramid of
  B-Spline functions, with control points distributed on grids of the following spacings:
  40$\times$100$\times$40 [mm] for the first (coarser) level, \isores[mm]{30} for the second level,
  and 20$\times$30$\times$10 [mm] in the third level.
Only the first and second levels included Gaussian smoothing of the target image ($\sigma$ = [2.0, 0.5] mm,
  respectively).
The actual choices of the parameter settings are publicly distributed with the source code for the
  experiments\footnote{\url{https://github.com/oesteban/RegSeg/tree/master/Scripts/pyacwereg/data/regseg_hcp.json}}.
These settings were obtained manually, driven by the feedback obtained from the post-registration
  convergence reports \citepalias[like that found in][section S1.3]{esteban_useful_2015}.
We released \regseg{} along with the tool to generate such convergence reports.

>>>>

\begin{figure}
  \includegraphics[width=\linewidth]{figure01}
  \caption{The {\color{blue} interfacing surfaces $\Gamma_m$ between the competing \glspl{roi} $\Omega_l$,
  play the role of active contours which drive the registration process.}
  They evolve iteratively following {\color{blue} the normal $\hat{\mathbf{n}}_i$ of the surface} at each vertex
    $\vec{v}_i$ of the mesh.
  The gradient speeds $\bar{s}_i$ drive registration, which are computed as the disparity of the data
    energies with respect to the two limiting regions of $M(\vec{v}_i)$, the features of the image
    $M$ in the location of vertex $\vec{v}_i$.
  {\color{blue} The computation of shape-gradients is developed in \ref{app:shape_gradients}.}
  In this figure, the $\bar{s}_1$ derived from \autoref{eq:regseg-shape_gradient_disc2} is 
    written in the lower box, with $\Omega_{wm}$ in the inner limiting region, $\Omega_{gm}$
    the outer region, and $w_{0,1}$ is the relative area associated with vertex $\vec{v}_1$
    with respect to the total area of surface $\Gamma_0$.
  }\label{fig:regseg-method}
\end{figure}

\begin{figure*}
% \includestandalone[width=\linewidth]{figure02-model}
\begin{minipage}{\textwidth}
\includegraphics[width=\linewidth]{figure02}\\%
\vskip-2.0em
\includegraphics[width=\linewidth]{legend}%x
\end{minipage}
% \includegraphics[width=\linewidth]{figure04B}
% \includegraphics[width=\linewidth]{figure04C}
\caption{%
Evolution of the segmentation model defined by the homogeneous regions $\Omega_l$,
  for one real dataset.
{\color{blue}%
Panel (A, left) shows the joint distribution of the \gls*{fa} and \gls*{adc} conditioned to
  the segmentation \omegaset{} defined by the surfaces $\gammaset_R$ extracted from
  the \gls*{t1} image.
The plot was generated for reference using undistorted diffusion data, and
  therefore, $\gammaset_R$ is aligned with the \gls*{fa} and the \gls*{adc}.
The problem arises when the diffusion data present deformation, and the
  contours $\gammaset_R$ do not fit within the data (A, center).
After registration with \regseg{}, the contours are mapped onto the
  diffusion data (A, right), and the joint density plot is closer to the reference
  situation.
In panel (B), the three plots in (A) are decomposed tissue-wise.
Using filled contours, the bivariate distribution of each tissue is highlighted in its designated color,
  and represented over the remaining tissues (in gray colors).
To help assessment, dashed contours in black-to-white colors represent
  the corresponding distribution in the reference plot.
The registration process optimizes the segmentation model of \regseg{},
  and thus, the distribution of each region after registration is located
  closer to that corresponding in the reference situation, the shape of the distribution
  is more similar to the reference, and their spread is also reduced.
The effects of optimization are more noticeable on the \gls*{gm} ($\Omega_\text{GM}$) and
  the \gls*{wm} ($\Omega_\text{WM}$).
Particularly, the \gls*{wm} typically shows a bimodal distribution when the contours
  \gammaset{} do not fit the data.}
% The left plot in each panel represents a kernel density estimate of the location and spread of
%   each region $\Omega_l$ in the bivariate feature space of $M$, which comprises
%   the \gls*{fa} and \gls*{adc} maps.
% At the top and the right margins, the marginal distributions of $\Omega_l$ on the \gls*{fa} and
%   \gls*{adc} are respectively displayed.
% On the right side, the sampled distribution of each region is shown in small scatter plot, in contrast to
%   the complete bivariate data.
% A) shows the initialization of the model, when the anatomically-correct surfaces are not
%   mapped into the \gls*{dmri} space.
% The panel shows the resulting conditional distributions of segmenting the (distorted) \gls*{fa} and
%   \gls*{adc} maps with the original (undistorted) surfaces.
% Panel B shows the equivalent information, using the surfaces that are perfectly mapped into the
%   distorted \gls*{dmri} space.
% Since the surfaces are now aligned in $M$, the spreads of the conditional distributions for each region are
%   reduced with respect to those in A.
% Panel C, after \regseg{} registration, shows how the covariance matrices $\boldsymbol{\Sigma}_l$
%   are minimized with respect to panel A, and the resulting segmentation is close
%   to the theoretical situation presented in panel B.}
}\label{fig:regseg-model}
\end{figure*}

@@

\revcomment[R\#3-C.20]{%
  \begin{align}
  \frac{\partial E(\vec{u})}{\partial \vec{u}_k} &=
  \frac{ \partial }{\partial \vec{u}_k} \Big\{
  \underset{l}{\sum} \big[ \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \big] \,
  + \int_{\Omega} \frac12 [ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
  + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} ] \,d\vec{r}
  \Big\},
  \label{eq:regseg-gradient_descent}
  \end{align}}
  where $\vec{u}^{\circ2} = \vec{u}^T \cdot \vec{u}$, \revcomment[R\#3-C.21]{%
  and $\{\boldsymbol{\alpha}, \boldsymbol{\beta}\}$ are the expected variances along
  the imaging axes of the displacements field and its gradient, respectively.}
Then, the data and regularization terms are split and discretized to compute their
  derivatives.
\revcomment[R\#3-C.14]{%
The derivative of the data term is computed using explicit shape gradients (see \ref{app:shape_gradients}),
  that finally lead to obtain vertex-wise speeds of the gradient $\bar{s}_i$ as}
  illustrated in \autoref{fig:regseg-method}.
\revcomment[R\#3-C.14]{%
Using the expressions \eqref{eq:regseg-gradient_wshape} and 
  \eqref{eq:regseg-shape_gradient_final} given in \ref{app:shape_gradients}
  and introducing the analytical derivative of the regularization term, then
  the \autoref{eq:regseg-gradient_descent} is reformulated as:}

  \begin{align}
  \frac{\partial E(\vec{u})}{\partial \vec{u}_k} &=
  \vec{g}_k  + \boldsymbol{\alpha} \cdot \vec{u}_k - \boldsymbol{\beta} \cdot (\Delta \vec{u}_k),
  \label{eq:regseg-final_gradient}
  \end{align}
\revcomment[R\#3-C22]{%
where $\vec{g}_k$ are the shape-gradient contributions on the coefficients $\vec{u}_k$ of the B-spline grid.}
Finally, to descend this gradient, we establish a semi-implicit Euler scheme
  \citepalias[see][section S1.3]{esteban_useful_2015},
  with a step size parameter $\delta$, which we solve in the spectral domain as follows:

@@

\paragraph*{Implementation details and convergence}%
\label{sec:regseg-conv_report}
The \regseg{} tool includes a multiresolution strategy on the free-form deformation field.
Registration pyramids are created by setting the spacing between the control points of the B-spline basis
  functions for each level of the multiresolution strategy.
\revcomment[R\#3-C32]{As a rule of thumb, for a $\delta =$ 1.0 both $\boldsymbol{\alpha}$
  and $\boldsymbol{\beta}$ will typically be in the range [0.0, 1.0].}
The implementation details as well as other features (such as the sparse matrix approach
  to fast interpolation) and the main parameters
  ($\delta$, $\boldsymbol{\alpha}$, $\boldsymbol{\beta}$, the B-spline grid resolutions,
  and target image smoothing) are discussed in the \citetalias[][section S1]{esteban_useful_2015}.

@@

\begin{figure*}
\includegraphics[width=\linewidth]{figure03}
\caption{Evaluation of \regseg{} using phantom data according to the following instrumental workflow.
  1) The reference surfaces $\gammaset_R$ are triangle meshes extracted from the four binary shapes (i.e., ``box'', ``ball'', ``L'', ``gyrus'').
  2) A ground-truth displacement field was generated as described in \autoref{sec:regseg-digital_phantoms}, and applied to warp
      $\gammaset_R$, thereby obtaining $\gammaset_\text{true}$.
  3) After being warped, $\gammaset_\text{true}$ were projected onto the corresponding discrete 3D volume and downsampled to create partial volume effects at two resolutions,
     i.e., \isores[mm]{2.0} and \isores[mm]{1.0}, thereby producing sets of tissue fractions maps.
  4) The tissue fractions were fed into a \acrfull*{mr} simulator, which generated \acrfull*{t1} and \acrfull*{t2} -like images at the
     two possible resolutions.
  5) The \regseg{} tool was applied using the warped test images as multispectral moving images and $\gammaset_R$ as shape priors.
  6) The agreement between the surfaces fitted by \regseg{} ($\hat{\gammaset}_\text{test}$) and $\gammaset_\text{true}$ were assessed
     visually using automatically generated visual reports and quantitatively with the Hausdorff distance between the
     corresponding surfaces.}\label{fig:regseg-evphantoms}
\end{figure*}

@@

\paragraph*{Settings} %
\Regseg{} accepts an affine mapping from surface-space to the \gls*{dmri} data as initialization.
However, the images provided by the \gls*{hcp} are already spatially normalized.
Therefore, the initial estimation of distortion is zero in this experiment.
Since the distortion $U_\text{true}$ is aligned along the \gls*{pe} direction ($y$-axis in our settings),
  \regseg{} was configured to allow nonzero displacements only on that corresponding direction.
For the experiments on real data, \regseg{} established a multi-resolution pyramid of
  B-spline functions, with control points distributed on grids of the following spacings:
  40$\times$100$\times$40 [mm] for the first (coarser) level, \isores[mm]{30} for the second level,
  and 20$\times$30$\times$10 [mm] in the third level.
Only the first and second levels included Gaussian smoothing of the target image ($\sigma$ = [2.0, 0.5] mm,
  respectively).
The actual choices of the parameter settings are publicly distributed with the source code for the
  experiments\footnote{\url{https://github.com/oesteban/RegSeg/tree/master/Scripts/pyacwereg/data/regseg_hcp.json}}.
These settings were obtained manually, driven by the feedback obtained from the post-registration
  convergence reports \citepalias[like that found in][section S1.3]{esteban_useful_2015}.
We released \regseg{} along with the tool to generate such convergence reports.

### 2016-02-02 Publications/2015-NeuroImage/05-results.tex
added fixes of all coauthors, and spell check

\begin{figure*}[!h]
  \centering
  \includestandalone[width=\linewidth]{figure05}
  \caption{A. Visual assessment of the results obtained on the {\color{blue} digital phantoms.
    The panel shows three coronal slices corresponding to the small numbers annotated on the lower
      left corner of each cut, for} the low resolution sets:
    ``gyrus'' (top left), ``L'' (top right), ``ball'' (bottom left),
    and ``box'' at (bottom right).
  The contours recovered after registration are represented in yellow.
  \Regseg{} achieved high accuracy because it determined the almost exact locations of the registered
    contours with respect to their ground truth positions (shown in green).
  The partial volume effect makes segmentation of the sulci a challenging problem with voxel-wise
    clustering methods, but they were successfully segmented with \regseg{}.
  B. Quantitative evaluation of registration errors in terms of the average Hausdorff distances between
    surfaces at low (left) and high (right) resolutions, which demonstrate that the errors were
    consistently below the size of the voxels.
    }\label{fig:regseg-phantom}
\end{figure*}
\begin{figure*}[!h]
  \centering
  \includestandalone[width=\linewidth]{figure06}
  \caption{A. Example of a visual assessment report, which was automatically generated by the evaluation tool.
    Each view shows one component of the input image (in this case, the \gls*{fa} map), the ground-truth locations
    of the surfaces (green contours), and the resulting surfaces obtained with the test method (yellow contours).
  The first two rows show axial slices for \regseg{} and the \acrfull*{t2b} method, while the last two rows
    show the corresponding sagittal views.
  The coronal view is omitted because it was the least informative due to the directional property
    of the distortions.
  Specific regions where \regseg{} outperformed \gls*{t2b} are enlarged.
  B. Violin plots of the error distributions for each surface {\color{blue} across the 16 subjects}, which show the 
    voxel size of the \gls*{dmri} images (1.25 mm), thereby supporting the improved results obtained
    by \regseg{} with the proposed settings.
  }\label{fig:regseg-results_real}
\end{figure*}

>>>>

\begin{figure*}[!h]
  \centering
  \includestandalone[width=\linewidth]{figure05}
  \caption{A. Visual assessment of the results obtained on the {\color{blue} digital phantoms.
    The panel shows three coronal slices (indices indicated on the lower left corner) of each phantom
    volume at low resolution}:
    ``gyrus'' (top left), ``L'' (top right), ``ball'' (bottom left),
    and ``box'' at (bottom right).
  The contours recovered after registration are represented in yellow.
  \Regseg{} achieved high accuracy because it determined the almost exact locations of the registered
    contours with respect to their ground truth positions (shown in green).
  The partial volume effect makes segmentation of the sulci a challenging problem with voxel-wise
    clustering methods, but they were successfully segmented with \regseg{}.
  B. Quantitative evaluation: {\color{blue} the violin plot shows the variability across experiments
    of the average Hausdorff distance measured in each vertex of the corresponding surface, for the}
    low (left) and high (right) resolutions. 
    {\color{blue} Error averages were} consistently below the size of the voxel.
    }\label{fig:regseg-phantom}
\end{figure*}
\begin{figure*}[!h]
  \centering
  \includestandalone[width=\linewidth]{figure06}
  \caption{A. Example of a visual assessment report, which was automatically generated by the evaluation tool.
    Each view shows one component of the input image (in this case, the \gls*{fa} map), the ground-truth locations
    of the surfaces (green contours), and the resulting surfaces obtained with the test method (yellow contours).
  The first two rows show axial slices for \regseg{} and the \acrfull*{t2b} method, while the last two rows
    show the corresponding sagittal views.
  The coronal view is omitted because it was the least informative due to the directional property
    of the distortions.
  Specific regions where \regseg{} outperformed \gls*{t2b} are enlarged.
  B. Violin plots of the error distributions for each surface {\color{blue} across the 16 subjects}, which show the 
    voxel size of the \gls*{dmri} images (1.25 mm), thereby supporting the improved results obtained
    by \regseg{} with the proposed settings.
  }\label{fig:regseg-results_real}
\end{figure*}

### 2016-02-02 Publications/2015-NeuroImage/06-discussion.tex
added fixes of all coauthors, and spell check

The proposed application of the method in the task of identifying structural information
  in \gls*{dmri} images is an active field of research \citep{jeurissen_tissuetype_2015}.
\revcomment[R\#3-C.9]{%
Current processing of \gls{dmri} involved in the connectome extraction and other applications
  (such as \gls*{tbss} or surgical planning)} require a precise segmentation
  of the anatomical structures in the diffusion space.
Some examples of these processing tasks are the structure-informed reconstruction of \gls*{dmri}
  data \citep{jeurissen_multitissue_2014,daducci_accelerated_2015}, the anatomically constrained
  tractography \citep{smith_anatomicallyconstrained_2012}, and the imposition of the cortical
  parcellation mapped from the \gls*{t1} image \citep{hagmann_mapping_2008}.
The problem was firstly addressed using image segmentation approaches in the native diffusion
  space, without definite and compelling results.
With the introduction of retrospective correction methods for the \emph{\gls*{epi} distortions}
  and image registration approaches, the task has been typically solved in a two-step approach.
First, the \glspl*{dwi} are corrected for \emph{\gls*{epi} distortions} by estimating
  the nonlinear deformation field from extra MR acquisitions
  \citep{jezzard_correction_1995,chiou_simple_2000,cordes_geometric_2000,kybic_unwarping_2000}.
Second, mapping the structural information from the corresponding \gls*{t1} image
  using a linear registration tool like \emph{bbregister} \citep{greve_accurate_2009}.
The current activity on improving correction methods \citep{irfanoglu_drbuddi_2015} and
  the comeback of segmentation of \gls*{dmri} in its native space
  \citep{jeurissen_tissuetype_2015} proof the open interest of this application.
\Regseg{} addresses this joint problem in a single step and it does not require any additional
  acquisition other than the minimal protocol comprehending only \gls*{t1} and \gls*{dmri} images.
This situation is found commonly in historical datasets.
\revcomment[R\#1-C.2]{%
We envision \regseg{} to be integrated in diffusion processing pipelines, after a preliminary
  \gls*{dti} computation and before anatomically-informed reconstruction and tractography
  methods.
Since the structural information is projected into the native space of \gls*{dmri},
  these two processes and the matrix building task can be performed on the unaltered
  \gls*{dmri} signal (i.e. without resampling data to an undistorted space).}
\revcomment[R\#3-C.9]{%
For analyses other than connectivity, like \gls*{tbss}, the deformation estimated by \regseg{}
  can be used to map the tracts into reference space.}

>>>>

The proposed application of the method in the task of identifying structural information
  in \gls*{dmri} images is an active field of research \citep{jeurissen_tissuetype_2015}.
\revcomment[R\#3-C.9]{%
Current processing of \gls{dmri} involved in the connectome extraction and other applications
  (such as \gls*{tbss} or surgical planning)} require a precise segmentation
  of the anatomical structures in the diffusion space.
Some examples of these processing tasks are the structure-informed reconstruction of \gls*{dmri}
  data \citep{jeurissen_multitissue_2014,daducci_accelerated_2015}, the anatomically constrained
  tractography \citep{smith_anatomicallyconstrained_2012}, and the imposition of the cortical
  parcellation mapped from the \gls*{t1} image \citep{hagmann_mapping_2008}.
The problem was firstly addressed using image segmentation approaches in the native diffusion
  space, without definite and compelling results.
With the introduction of retrospective correction methods for the \emph{\gls*{epi} distortions}
  and image registration approaches, the task has been typically solved in a two-step approach.
First, the \glspl*{dwi} are corrected for \emph{\gls*{epi} distortions} by estimating
  the nonlinear deformation field from extra MR acquisitions
  \citep{jezzard_correction_1995,chiou_simple_2000,cordes_geometric_2000,kybic_unwarping_2000}.
Second, mapping the structural information from the corresponding \gls*{t1} image
  using a linear registration tool like \emph{bbregister} \citep{greve_accurate_2009}.
The current activity on improving correction methods \citep{irfanoglu_drbuddi_2015} and
  the comeback of segmentation of \gls*{dmri} in its native space
  \citep{jeurissen_tissuetype_2015} proof the open interest of this application.
\Regseg{} addresses this joint problem in a single step and it does not require any additional
  acquisition other than the minimal protocol comprehending only \gls*{t1} and \gls*{dmri} images.
This situation is found commonly in historical datasets.
\revcomment[R\#1-C.2]{%
We envision \regseg{} to be integrated in diffusion processing pipelines, after a preliminary
  \gls*{dti} computation and before anatomically-informed reconstruction and tractography
  methods.
Since the structural information is projected into the native space of \gls*{dmri},
  these two processes and the matrix building task can be performed on the unaltered
  \gls*{dmri} signal (i.e. without resampling data to an undistorted space).}
\revcomment[R\#3-C.9]{%
For analyses other than connectivity, like \gls*{tbss}, the deformation estimated by \regseg{}
  can be used to map the tracts into structural space.}

### 2016-02-02 Publications/2015-NeuroImage/08-appendix-00-shapepriors.tex
added fixes of all coauthors, and spell check

\revcomment[R\#3-C.22]{%
Then, we identify the shape gradient contribution $\vec{g}_k$ on the coefficients $\vec{u}_k$ of the B-Spline grid:
  \begin{equation}
  \label{eq:regseg-gradient_wshape}
  \begin{split}
  \vec{g}_k &= - \underset{i}{\sum} \left\langle \frac{\partial \vec{v}_i'}{\partial \vec{u}_k}, \bar{s}_i'\right\rangle \\
  \text{with }
  \bar{s}_i' &= w_i \left[ \mdist{f_i'}{out} - \mdist{f_i'}{in} \right] \, \hat{\vec{n}}_i, \\
  \text{and }
  \frac{\partial \vec{v}_i'}{\partial \vec{u}_k} &=
  \frac{\partial}{\partial \vec{u}_k} \left\{ \vec{v}_i + \sum_k \psi_k(\vec{v}_i) \vec{u}_k \right\} = \psi_k(\vec{v}_i)\, \hat{\vec{e}},
  \end{split}
  \end{equation}%
  where $\hat{\vec{e}}$ is the coordinates system's unit vector.
Therefore, the shape gradients projected to the grid of B-Spline control points is:
\begin{equation}
  \vec{g}_k = - \underset{i}{\sum} \bar{s}_i \cdot \psi_k(\vec{v}_i) \, \hat{\vec{e}}.
  \label{eq:regseg-shape_gradient_final}
\end{equation}}

@@

In general, the anisotropy \revcomment[R\#3-C.31]{of the distortion field} is aligned with the 
  \revcomment[R\#3-C.31]{voxel coordinate system}, so
  $\mathbf{A}$ and $\mathbf{B}$ of \eqref{eq:regseg-energy} can be simplified to diagonal matrices
  \revcomment[R\#3-C.31]{to regularize the regitration process}, such that
  $\mathbf{A}= \, \boldsymbol{\alpha}\,\vec{I}_n$ and
  $\mathbf{B}= \, \boldsymbol{\beta}\,\vec{I}_n$.
By substituting into equation \eqref{eq:regseg-energy}, we obtain:

>>>>

\revcomment[R\#3-C.22]{%
Then, we identify the shape gradient contribution $\vec{g}_k$ on the coefficients $\vec{u}_k$ of the B-spline grid:
  \begin{equation}
  \label{eq:regseg-gradient_wshape}
  \begin{split}
  \vec{g}_k &= - \underset{i}{\sum} \left\langle \frac{\partial \vec{v}_i'}{\partial \vec{u}_k}, \bar{s}_i'\right\rangle \\
  \text{with }
  \bar{s}_i' &= w_i \left[ \mdist{f_i'}{out} - \mdist{f_i'}{in} \right] \, \hat{\vec{n}}_i, \\
  \text{and }
  \frac{\partial \vec{v}_i'}{\partial \vec{u}_k} &=
  \frac{\partial}{\partial \vec{u}_k} \left\{ \vec{v}_i + \sum_k \psi_k(\vec{v}_i) \vec{u}_k \right\} = \psi_k(\vec{v}_i)\, \hat{\vec{e}},
  \end{split}
  \end{equation}%
  where $\hat{\vec{e}}$ is the coordinates system's unit vector.
Therefore, the shape gradients projected to the grid of B-spline control points is:
\begin{equation}
  \vec{g}_k = - \underset{i}{\sum} \bar{s}_i \cdot \psi_k(\vec{v}_i) \, \hat{\vec{e}}.
  \label{eq:regseg-shape_gradient_final}
\end{equation}}

@@

In general, the anisotropy \revcomment[R\#3-C.31]{of the distortion field} is aligned with the 
  \revcomment[R\#3-C.31]{voxel coordinate system}, so
  $\mathbf{A}$ and $\mathbf{B}$ of \eqref{eq:regseg-energy} can be simplified to diagonal matrices
  \revcomment[R\#3-C.31]{to regularize the registration process}, such that
  $\mathbf{A}= \, \boldsymbol{\alpha}\,\vec{I}_n$ and
  $\mathbf{B}= \, \boldsymbol{\beta}\,\vec{I}_n$.
By substituting into equation \eqref{eq:regseg-energy}, we obtain:

### 2016-02-02 Publications/2015-NeuroImage/08-appendix-01.tex
added fixes of all coauthors, and spell check

In the probabilistic approach to image segmentation, the smoothness constraints
can be introduced into the \gls{map} criterium as a \gls{mrf}. Accordingly to the
Hammersley-Clifford theorem (\textbf{FIXME}: citations needed), an \gls{mrf} can
be equivalently characterized by a Gibb's distribution:
\begin{equation}
P(\mathbf{K})=\underset{i}{\prod} Z(U)^{-1}\,e^{-U(\mathbf{x}_i \mid \beta,K)} = \mathit{const.}\,e^{- \sum\limits_i \left( U(\mathbf{x}_i \mid \beta,K) \right) } \sim e^{-\nu_B \left| K \right| }.
\end{equation}
This way, we include the assumption that the total length of the edge set $K$ is small
and we draw the equivalence with the \gls{mrf} modeling, for $U(\mathbf{x}_i \mid \beta,K)$
being the Potts model.

In the probabilistic approach, the underlying concept to segmenting an image is the
Bayes' rule. For a discrete image of $N$ pixels indexed by $i$, this rule reads as
follows:
\begin{equation}
P(K|I)=P(I \mid K)\, P(K)=\underset{i}{\prod} p (\mathbf{x}_i | K ) \, p_i(K).
\end{equation}
The normalizer probability $P(I)$ has been ommited for simplicity.

>>>>

In the probabilistic approach to image segmentation, the smoothness constraints
can be introduced into the \gls{map} criterion as a \gls{mrf}. Accordingly to the
Hammersley-Clifford theorem (\textbf{FIXME}: citations needed), an \gls{mrf} can
be equivalently characterized by a Gibb's distribution:
\begin{equation}
P(\mathbf{K})=\underset{i}{\prod} Z(U)^{-1}\,e^{-U(\mathbf{x}_i \mid \beta,K)} = \mathit{const.}\,e^{- \sum\limits_i \left( U(\mathbf{x}_i \mid \beta,K) \right) } \sim e^{-\nu_B \left| K \right| }.
\end{equation}
This way, we include the assumption that the total length of the edge set $K$ is small
and we draw the equivalence with the \gls{mrf} modeling, for $U(\mathbf{x}_i \mid \beta,K)$
being the Potts model.

In the probabilistic approach, the underlying concept to segmenting an image is the
Bayes' rule. For a discrete image of $N$ pixels indexed by $i$, this rule reads as
follows:
\begin{equation}
P(K|I)=P(I \mid K)\, P(K)=\underset{i}{\prod} p (\mathbf{x}_i | K ) \, p_i(K).
\end{equation}
The normalizer probability $P(I)$ has been omitted for simplicity.
